---
title: "Olympic Games"
format: 
  dashboard:
    orientation: columns
    nav-buttons: [github]
    github: https://github.com/posit-conf-2024/olympicdash
logo: images/olympics-logo.svg
logo-alt: "Olympics logo with multicolored circles."
---

```{r}
#| label: load-packages
library(tidyverse)
library(gt)
library(nord)
library(paletteer)
```

```{r}
#| label: load-data
olympics_full <- read_csv("data/olympics.csv")
```

```{r}
#| label: prep-data
olympics <- olympics_full |>
  filter(
    !is.na(medal)
  ) |>
  separate_wider_delim(
    cols = team, 
    names = c("team", "suffix"), 
    delim = "-", 
    too_many = "merge", 
    too_few = "align_start"
  ) |>
  select(-suffix) |>
  mutate(medal = fct_relevel(medal, "Bronze", "Silver", "Gold"))
```

# Summer Olympics

```{r}
#| label: summer-prep-data
summer_olympics <- olympics |> 
  filter(season == "Summer")

```

## Column - Medals by sport and year {width="65%"}

### Row - Medals by sport {height="60%"}

```{r}
#| label: medals-by-sport
#| title: Medals by sport
#| fig-width: 10
#| fig-asp: 0.618
summer_olympics |>
  mutate(
    sport = fct_lump_n(sport, n = 15),
    sport = fct_infreq(sport),
    sport = fct_rev(sport),
    sport = fct_relevel(sport, "Other", after = 0)
  ) |>
  ggplot(aes(y = sport, fill = medal)) +
  geom_bar() +
  guides(fill = guide_legend(reverse = TRUE)) +
  labs(
    x = NULL,
    y = NULL,
    fill = "Medal"
  ) +
  theme_minimal() +
  theme(
    legend.position = "inside",
    legend.position.inside = c(0.9, 0.2),
    legend.direction = "horizontal",
    legend.background = element_rect(fill = "white", color = "gray")
  )
```

### Row - Medals by year {height="40%"}

::: {.card title="Medals by Year"}
Due to World War II, no Olympic Games were held in 1940 and 1944.

```{r}
#| label: medals-by-year
#| fig-asp: 0.3
#| fig-width: 10
summer_olympics |>
  count(year, medal) |>
  ggplot(aes(x = year, y = n, color = medal)) +
  geom_point(size = 0.5) +
  geom_line() +
  guides(color = guide_legend(reverse = TRUE)) +
  scale_x_continuous(breaks = seq(1896, 2020, 8)) +
  labs(
    x = "Year",
    y = NULL,
    color = "Medal"
  ) +
  theme_minimal() +
  theme(
    legend.position = "inside",
    legend.position.inside = c(0.9, 0.2),
    legend.direction = "horizontal",
    legend.background = element_rect(fill = "white", color = "gray")
  )
```
:::

## Column - Medals by country {width="35%"}

### Row - Value Boxes {height="30%}

```{r}
#| label: summer-calculate-most-medals

summer_most_golds<- summer_olympics |> 
  filter(medal == "Gold") |> 
  count(team, sort = TRUE) |> 
  slice_head(n = 1)

summer_most_silvers<- summer_olympics |> 
  filter(medal == "Silver") |> 
  count(team, sort = TRUE) |> 
  slice_head(n = 1)

summer_most_bronzes<- summer_olympics |> 
  filter(medal == "Bronze") |> 
  count(team, sort = TRUE) |> 
  slice_head(n = 1)

```

:::{.valuebox icon="award-fill" color="#d4ad37"}
Most golds:

`{r} summer_most_golds$n`

`{r} summer_most_golds$team`

:::

:::{.valuebox icon="award-fill" color="#c0c0c0"}
Most silvers:

`{r} summer_most_silvers$n`

`{r} summer_most_silvers$team`

:::

:::{.valuebox icon="award-fill" color="#cd7f32"}
Most bronzes:

`{r} summer_most_bronzes$n`

`{r} summer_most_bronzes$team`

:::

### Row - Tabsets of Tables {height="70% .tabset}

::: {.card title="Top 30 Medals"}
Teams ordered in descending order of medals

```{r}
#| label: top-30-medals
summer_olympics |>
  count(team, medal) |>
  pivot_wider(
    names_from = medal, 
    values_from = n, 
    values_fill = 0
  ) |>
  mutate(Total = Bronze + Gold + Silver) |>
  relocate(Team = team, Gold, Silver, Bronze, Total) |>
  arrange(desc(Total), Team) |>
  slice_head(n = 30) |> 
  gt() |> 
  cols_align(align = "left", column = "Team") |> 
    data_color(
    method = "numeric",
    palette = "nord::aurora"
  )
```
:::

::: {.card title="Bottom 30 medals"}

Teams ordered in descending order of medals

```{r}
#| label: bottom-30-medals
summer_olympics |>
  count(team, medal) |>
  pivot_wider(
    names_from = medal, 
    values_from = n, 
    values_fill = 0
  ) |>
  mutate(Total = Bronze + Gold + Silver) |>
  relocate(Team = team, Gold, Silver, Bronze, Total) |>
  arrange(Total, Team) |>
  slice_head(n = 30) |> 
  gt() |> 
  cols_align(align = "left", column = "Team") |> 
  data_color(
    method = "numeric",
    palette = "nord::aurora"
  )
```
:::

# Winter Olympics

```{r}
#| label: winter-prep-data
winter_olympics <- olympics |> 
  filter(season == "Winter")

```

## Column - Medals by sport and year {width="65%"}

### Row - Medals by sport {height="60%"}

```{r}
#| label: winter-medals-by-sport
#| title: Medals by sport
#| fig-width: 10
#| fig-asp: 0.618
winter_olympics |>
  mutate(
    sport = fct_lump_n(sport, n = 15),
    sport = fct_infreq(sport),
    sport = fct_rev(sport),
    sport = fct_relevel(sport, "Other", after = 0)
  ) |>
  ggplot(aes(y = sport, fill = medal)) +
  geom_bar() +
  guides(fill = guide_legend(reverse = TRUE)) +
  labs(
    x = NULL,
    y = NULL,
    fill = "Medal"
  ) +
  theme_minimal() +
  theme(
    legend.position = "inside",
    legend.position.inside = c(0.9, 0.2),
    legend.direction = "horizontal",
    legend.background = element_rect(fill = "white", color = "gray")
  )
```

### Row - Medals by year {height="40%"}

::: {.card title="Medals by Year"}
Due to World War II, no Olympic Games were held in 1940 and 1944.

```{r}
#| label: winter-medals-by-year
#| fig-asp: 0.3
#| fig-width: 10
winter_olympics |>
  count(year, medal) |>
  ggplot(aes(x = year, y = n, color = medal)) +
  geom_point(size = 0.5) +
  geom_line() +
  guides(color = guide_legend(reverse = TRUE)) +
  scale_x_continuous(breaks = seq(1896, 2020, 8)) +
  labs(
    x = "Year",
    y = NULL,
    color = "Medal"
  ) +
  theme_minimal() +
  theme(
    legend.position = "inside",
    legend.position.inside = c(0.9, 0.2),
    legend.direction = "horizontal",
    legend.background = element_rect(fill = "white", color = "gray")
  )
```
:::

## Column - Medals by country {width="35%"}

### Row - Value Boxes {height="30%}

```{r}
#| label: winter-calculate-most-medals

winter_most_golds<- winter_olympics |> 
  filter(medal == "Gold") |> 
  count(team, sort = TRUE) |> 
  slice_head(n = 1)

winter_most_silvers<- winter_olympics |> 
  filter(medal == "Silver") |> 
  count(team, sort = TRUE) |> 
  slice_head(n = 1)

winter_most_bronzes<- winter_olympics |> 
  filter(medal == "Bronze") |> 
  count(team, sort = TRUE) |> 
  slice_head(n = 1)

```

:::{.valuebox icon="award-fill" color="#d4ad37"}
Most golds:

`{r} winter_most_golds$n`

`{r} winter_most_golds$team`

:::

:::{.valuebox icon="award-fill" color="#c0c0c0"}
Most silvers:

`{r} winter_most_silvers$n`

`{r} winter_most_silvers$team`

:::

:::{.valuebox icon="award-fill" color="#cd7f32"}
Most bronzes:

`{r} winter_most_bronzes$n`

`{r} winter_most_bronzes$team`

:::

### Row - Tabsets of Tables {height="70% .tabset}

::: {.card title="Top 30 Medals"}
Teams ordered in descending order of medals

```{r}
#| label: winter-top-30-medals
winter_olympics |>
  count(team, medal) |>
  pivot_wider(
    names_from = medal, 
    values_from = n, 
    values_fill = 0
  ) |>
  mutate(Total = Bronze + Gold + Silver) |>
  relocate(Team = team, Gold, Silver, Bronze, Total) |>
  arrange(desc(Total), Team) |>
  slice_head(n = 30) |> 
  gt() |> 
  cols_align(align = "left", column = "Team") |> 
    data_color(
    method = "numeric",
    palette = "nord::aurora"
  )
```
:::

::: {.card title="Bottom 30 medals"}

Teams ordered in descending order of medals

```{r}
#| label: winter-bottom-30-medals
winter_olympics |>
  count(team, medal) |>
  pivot_wider(
    names_from = medal, 
    values_from = n, 
    values_fill = 0
  ) |>
  mutate(Total = Bronze + Gold + Silver) |>
  relocate(Team = team, Gold, Silver, Bronze, Total) |>
  arrange(Total, Team) |>
  slice_head(n = 30) |> 
  gt() |> 
  cols_align(align = "left", column = "Team") |> 
  data_color(
    method = "numeric",
    palette = "nord::aurora"
  )
```
:::
